---
title: "Interventions"
output:
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
    number_sections: true
    smart: false
---

```{r setup, include = FALSE}

```

<!-- WARN: auto order scenarios! -->


<!-- TODO: make scenarios similar to the best_guess ones. (based on previous) -->

# General

- PrEP modalities:
   - clinical: I use "clinical", "standard" or "STD" PrEP interchangeably.
   - OTC PrEP: Over The Counter PrEP
   - Any: used to count the number of individuals on any type of PrEP, they are
     exclusive so it's STD + OTC
- In scenarios without STD PrEP, no one starts STD PrEP anymore but the current
  STD PrEP users stay on PrEP until they "naturally" discontinue (random or
  lapse in indications).
- HBV: flares rate are always the ones from the litterature (30%)
	- don't interpret them the same way
	- STD PrEP: probably managed
	- OTC PrEP: probably not managed
- I often mention "rate based testing" and "interval testing"
    - A rate based testing is a probability `p` of being tested at every time
      step. (coin flip)
    - An interval testing is a deterministic process where the next test occurs
      after a number of steps defined by the interval. This approach can lead to
      immediate tests for GFR when starting PrEP for the first time as no GFR
      test was ever performed before.

**These are preliminary results:** The models were run with only 32 replications
to get a quick glance at the effects. The final results will use ~500 to 1000
runs per scenarios

## Outcomes names

- IR100: Incidence for 100 person year at risk
- IR100k: Incidence for 100k person year at risk
- NIA: Number of Infection Averted over 10 years (uses `baseline` as reference)
- PIA: Percentage of Infection Averted
- (10y): outcome calculated over the 10 years of intervention
- (ly): outcome calculated over the last year of intervention
- PrEP coverages are for each modalities $/frac{users}{indicated}$

<!-- TODO: hbv flare: ir100 out of all HBV infected and not `num` -->

## Quirks

"Clinical PrEP Mean Number of Unique Episodes (ly)" also takes into account the
PrEP episodes before the start of intervention. So it will often be higher than
for OTC.

# Scenarios

Scenarios start from a population "at equilibrium" which means that most
demographic indicators would remain stable over time if no changed was made.

Each scenario runs for 10 years.

## Generic Scenarios

These scenarios are to explore the model in a more "technical" way. To see what
is affected when a parameter is changed and when extreme values are used.

When a parameter is not described for a scenario it means that its value is
similar to the one in `baseline`.

### `baseline` - Reference Scenario

Reference scenario. What happened if everything stayed at equilibrium after
calibration.

- Default values
- No OTC PrEP

### `no_otc_prep_or_xx` - Increasing STD PrEP Coverage

Similar to `baseline` but with increased PrEP coverage. We get the increased
coverage by increasing the `prep.start.rate` by an odds ratio of 1.5 or 2. The
other PrEP parameters are kept the same (discontinuation, indications, etc).

### `only_otc_same_1` - Replace STD PrEP with OTC Same

Turns off STD PrEP and replace it with OTC PrEP. OTC PrEP here behaves exactly
like STD PrEP.

This is mostly a debugging scenario. It should yield very similar result to
`baseline`.

### `otc_same_xx` - Addition of OTC PrEP - Same as STD

PrEP OTC is turned on and behave exactly like STD PrEP (indication, gfr checks,
STI and HIV testing, etc).

The `prep.otc.start.rate` is based on `prep.start.rate` with an Odds Ratio
applied to it. So `otc_same_05` uses an odds ratio of 0.5 and `otc_same_15` an
odds ratio of 1.5.


### `otc_relaxed_xx` - Addition of OTC PrEP - Relaxed Indications

Similar to `otc_same_xx` but with relaxed OTC PrEP indications. In these
scenarios any individual sexually active (any sex within 6 month) without and
HIV positive diagnostic is eligible to OTC PrEP. The eligibility status is
tested every 6 months similar to STD PrEP.

The other PrEP attributes are kept similar to STD PrEP.

This raises the number of PrEP OTC eligibles from 48k to 60k (+25%).


### `only_otc_relaxed_base_xx` - Only OTC Relaxed

Exaclty the same as `otc_same_xx` but with STD PrEP turned off.

Use these scenarios to find the correct OR leading to the same number of PrEP
users as `baseline` (it's `only_otc_relaxed_base_06`, so an OR of 2/3).
Later `only_otc_relaxed_****` scenarios all use this 2/3 OR for `prep.otc.start.rate`.

### `only_otc_relaxed_disc_` - Only OTC Relaxed With Changed Discontinuation

Similar to `only_otc_relaxed_base_06` but with changes in the discontinuation
intervals. Discontinuation intervals are multiplied by the values in the name:
`only_otc_relaxed_disc_075` have the discontinuation intervals multiplied by
0.75 (reduced) and `only_otc_relaxed_disc_150` have them multiplied by 1.5
(increased).

### `only_otc_relaxed_gfr_int` - Only OTC Relaxed With Changed GFR tests

<!-- TODO: use rate based gfr testing for these scenarios -->

Similar to `only_otc_relaxed_base_06` but with changes in the interval testings
for GFR. GFR is tested every 1, 2, 5 years or never (Inf).

Note that in these scenarios the testing is interval based. It means that
someone started OTC PrEP for the first time will always be tested at that time
and then after "interval". Another approach is available. See the `best_`
scenarios.

### `only_otc_relaxed_hiv_tst_ints` - Only OTC Relaxed With Changed HIV tests

<!-- TODO: use rate based intervals here -->

Similar to `only_otc_relaxed_base_06` but with no automatic HIV and STI testing
upon starts. Then the test interval for HIV is modified: 13, 26 and 52 weeks or
3, 6 and 12 months.

These are interval testing based on when the last HIV test was performed (even
when not on PrEP)

### `only_otc_relaxed_sti_screen` - Only OTC Relaxed With Changed STI tests

Similar to `only_otc_relaxed_hiv_tst_ints` but with changes to the STI screening
rates instead. The rates are modified by the OR in the names.

### `only_otc_relaxed_adhr` - Only OTC Relaxed With Changed Adherence

Similar to `only_otc_relaxed_sti_screen` but with changes to the adherence
distributions. "m20" is "minus 20%" and "p10" is "plus 10%". That is reduce the
highest adherence group by 20 percentage point or increase it by 10. The
ration between the 2 other adherence groups is kept the same.

### `only_otc_relaxed_switch2std_` - Only OTC Relaxed With Switch to STD

Similar to `only_otc_relaxed_base_06` but with a percentage of the OTC PrEP user
switching to STD PrEP each year. `only_otc_relaxed_switch2std_05` is a rate of
5% of OTC PrEP users would switch to STD. Note that no other way to start STD
PrEP is offered in these scenarios. Even for someone who previously was on STD.
He would have to start OTC again and switch to STD again.

## Best Guess Scenarios and variations

These scenarios are here to explore a more "realistic" future that could take
place if OTC PrEP was approved Tomorrow.

### `otc_best_guess` - Best Guess Scenario

- Total number of people on any kind of PrEP ~40% more that `baseline`
- Same start rate for OTC and STD PrEP: OR = 0.6 compared to `baseline`'s
  `prep.start.rate`
- Relaxed Indications for OTC PrEP:
  - sexually active (any sex within 6 month) without and HIV positive diagnostic
  - checked every time step
- Same adherence distribution
- Discontinuation interval increased by a factor 1.5 for OTC PrEP
- STI and HIV screening every 6 months (no test on start OTC PrEP)
  - interval based for HIV
  - rate for STI
- GFR test every 2 years
  - different from the `only_otc_relaxed_gfr_int` scenarios
  - rate based GFR test while on OTC PrEP

All `best_****` scenarios start from this base.

### `best_disc_` - Best Guess Scenario - Vary Discontinuation

Similar to `otc_best_guess` but apply a multiplying factor to the
discontinuation intervals.

### `best_gfr_same_int` - Best Guess Scenario - Same GFR

Similar to `otc_best_guess` but checks GFR every 6 months for high risk (age >
50 or last know GFR < 90) and 12 months otherwise. This still use the rate based
testing.

### `best_gfr_int_` - Best Guess Scenario - Vary GFR intervals

Similar to `otc_best_guess` but checks GFR every 1, 3 or 5. This still use the rate based
testing.

### `best_tests_int_` - Best Guess Scenario - Vary Tests Intervals

Similar to `otc_best_guess` but test for HIV and STI every 4, 12 of 24 months.
Not that HIV tests are still interval based and STI tests rate based.

### `best_adhr_` - Best Guess Scenario - Vary Adherence

Similar to `otc_best_guess` but vary OTC PrEP adherence distribution similarly
to the `only_otc_relaxed__adhr_` scenarios.

# Results

The results table contains a lot of columns, most of them are debugging outputs
used to check that the model behaved appropriatly. In the next section I will
focus on the results that seems relevant to me for each of the scenario groups.

## Generic Scenarios

### `baseline` - Reference Scenario

This scenario represent the equilibrium state where the model simply runs for 10
years without any changes. It's the baseline to compare the other models to.

### `no_otc_prep_or_xx` - Increasing STD PrEP Coverage

These scenarios introduce the effects and side effects of PrEP:

- Higher Any / STD PrEP number and coverage
- reduced HIV incidence and prevalence
- increased proportion of HIV+ diagnosed due to higher test rates while on PrEP
- lower STI (gono, chla) incidence due to higher STI screening rates
- reduction of the number of indicated probably due to reduced STI incidences
  (not HIV diag as OTC indic stays the same)
- increased GFR drop due to PrEP, but same proportion of PrEP user with GFR <
  60, meaning that the screening for GFR correctly stops PrEP when a drop is
  present.
- proportional increase in HBV flare
- proportional increase in treatment resistance

### `only_otc_same_1` - Replace STD PrEP with OTC Same

Similar enough to `baseline` to think it works correctly. A few individuals
remains on STD PrEP throughout the intervention. Meaning they were on it when it
starts and never discontinued.

### `otc_same_xx` - Addition of OTC PrEP - Same as STD

This is very similar to the `no_otc_prep_or_xx` scenarios as here OTC PrEP has
the same effect as STD PrEP. Looking at the `Any PrEP Number (ly)` column (H) we
see that 0.5 and 1.0 respectively give the same total number of PrEP user as the
1.5 and 2.0 `no_otc_prep_or_xx` scenarios.

### `otc_relaxed_xx` - Addition of OTC PrEP - Relaxed Indications

With these relaxed indications, the number of person on PrEP increases very
fast. Because all the services linked to STD PrEP also applies to OTC PrEP, the
same side effects are seen here.

Slightly less benefits on HIV compared to the same number of Any PrEP users
in the `no_otc_prep_or_xx` scenarios. It would make sense as the indicated should be slightly less at
risk. *Does that stay true with higher number of replications?*

PrEP user with GFR < 60 is slightly higher for OTC PrEP user, 2.9% instead of
2.6% on the higher scenarios. I suspect it's purely because of of the rate of
(re)starting can lead to longer period without tests.


### `only_otc_relaxed_base_xx` - Only OTC Relaxed

Seems to need slightly more "Any PrEP" to get the same benefits on HIV as
`baseline`. It would make sense as the indicated should be slightly less at
risk.

<!-- TODO: make a scenarios with the exact same coverage to check that. -->

Appart from that everything seems as expected (similar when same number of any
prep)

### `only_otc_relaxed_disc_` - Only OTC Relaxed With Changed Discontinuation

`only_otc_relaxed_base_06` would be the same as `only_otc_relaxed_disc_100` (OR
of 1, not in the scenarios).

Most outcomes are related to the coverage of PrEP. So increasing by reducing
discontinuation has the same effect as increasing start rate.

More OTC PrEP with GFR<60. I suspect it's because they stay longer on OTC PrEP
but not long enough to trigger the next GFR test.

The faster the discontinuation the more HBV flare could occur as they are tied
only to discontinuation.

### `only_otc_relaxed_gfr_int` - Only OTC Relaxed With Changed GFR tests

`only_otc_relaxed_base_06` would correspond to the scenario with same GFR checks
as STD (6 months for high risk and 12 months for low risk).

As expected, no more GFR drops due to PrEP compared to
`only_otc_relaxed_base_06` but higher proportion of OTC PrEP with GFR<60 when
testing is less frequent.

### `only_otc_relaxed_hiv_tst_ints` - Only OTC Relaxed With Changed HIV tests

`only_otc_relaxed_base_06` and `only_otc_relaxed_hiv_tst_ints` are the same.

Changes are too small in size to be visible. Note that the 2 years interval is
similar to the test rate in general population.

### `only_otc_relaxed_sti_screen` - Only OTC Relaxed With Changed STI tests

The ORs indeed reduce the test rate and increase the IR100 of gono and chla but
they are very noisy outcomes and more repetitions are required to assess
correctly.

### `only_otc_relaxed_adhr` - Only OTC Relaxed With Changed Adherence

`only_otc_relaxed_base_06` would correspond to the "no change" in adherence.

The effect is barely visible with the current number of replications. Either on
the HIV IR100 and the number of resistances over 10 years.

### `only_otc_relaxed_switch2std_` - Only OTC Relaxed With Switch to STD

Not much to say as so few people switch and stay on STD PrEP. More complex
scenarios would need to be devised to have something interesting to say.

## Best Guess Scenarios and variations

<!-- TODO: I should calc the NIA / PIA for these based on the best guess  -->
<!-- and not baseline -->

Evolution of STD vs OTC PrEP over time in best guess

```{r}
source("R/shared_variables.R", local = TRUE)

library(dplyr)
library(tidyr)
library(ggplot2)

theme_set(theme_light())

d <- readRDS("./data/run/scenarios/merged_tibbles/df__otc_best_guess.rds")

d |>
  select(sim, time, prepCurr, prep.otcCurr) |>
  mutate(prep_any = prepCurr + prep.otcCurr) |>
  pivot_longer(-c(sim, time)) |>
  ggplot(aes(x = time / 52, y = value, col = name)) +
  geom_smooth()
```

### `otc_best_guess` - Best Guess Scenario

Compared to Baseline:
- ~40% more PrEP users total
    - STD PrEP number is divided by 2 (6k instead of 12k)
    - OTC PrEP number is 11k
- HIV:
    - 11% PIA
    - prop diag: 88% vs 84%
- STI:
    - no visible effect
    - need more replications to validate that
- GFR:
    - higher number of drop (directly linked to Any PrEP number)
    - 6.3% PrEP users with GFR<60 vs 2.5%
    - 8% for PrEP OTC and 2.5% for STD PrEP
- HBV:
    - Same number on both regimen. Because less discont on OTC
    - if all OTC flare are unmanaged: 70 over 10year, IR100k: 5.1
    - still very low
- Resistance:
    - ~40% more: the number of additional PrEP users
    - 90 vs 65 over 10 years
    - very low

Conclusion here:
- desirable outcome
- GFR must be checked: communication is enough?

#### PrEP repartition by age



```{r include = FALSE, eval = FALSE}
sim <- readRDS("./data/run/scenarios/sim__otc_best_guess__1.rds")
age.breaks <- sim$param$netstats$demog$age.breaks
age_grp_names <- cut(age.breaks[-1], age.breaks) |>
  levels()
attr <- sim$run[[1]]$attr
prep_std <- attr$prep
prep_otc <- attr$prep.otc
age_grp <- attr$age.grp
```

STD PrEP

```{r eval = FALSE}
prop <- table(age_grp[prep_std == 1]) |>
  prop.table() |>
  (\(x) paste0(round(x * 100, 1), "%"))()
names(prop) <- age_grp_names
prop

#> (15,25] (25,35] (35,45] (45,55] (55,65]
#>  "24%" "25.8%" "22.3%" "16.8%" "11.2%"
```

OTC PrEP

```{r eval = FALSE}
prop <- table(age_grp[prep_otc == 1]) |>
  prop.table() |>
  (\(x) paste0(round(x * 100, 1), "%"))()
names(prop) <- age_grp_names
prop

#> (15,25] (25,35] (35,45] (45,55] (55,65]
#> "20.2%" "25.7%" "23.1%" "18.5%" "12.6%"
```

### `best_disc_` - Best Guess Scenario - Vary Discontinuation

Major impact on HIV outcomes, but still better than baseline in most case.

### `best_gfr_same_int` and `best_gfr_int_` - Vary GFR test ints

Better intervals reduces the amount of PrEP users with GFR<60. `same` is not as
good as STD PrEP as it does not automatically test upon OTC PrEP start.

Testing every year instead of 2 years without distinction in risk is still
better as well.

Note that higher GFR test rate slightly reduces PrEP coverage as it stop PrEP
for these users. In reality a change in regimen or move to STD PrEP could be
possible.

### `best_tests_int_` - Best Guess Scenario - Vary Tests Intervals

As with the generic scenarios on tests, the STI screening seems to be the major
driving force here. With HIV diagnosed proportion not varying much but IR100 for
STI being affected way more.

### `best_adhr_` - Best Guess Scenario - Vary Adherence

As with the generic scenario: not much of a visible effect. Requires more
replications and probably still not huge.

# Next Steps

## Tx resistance paper

https://academic.oup.com/cid/article/78/5/1240/7425528
-> does not change our estimates
-> could be very interesting to model resistance acquisition with "recent" use and not only while using
-> PrEP use & resistance in Dat'Aids?

## Calibration

Not conviced by the current PrEP calibration (older values).
-> as Christina after her defense
Need to recalibrate

## Replications

Run 500-1000 replications

## Scenarios

- Better values for the `only_otc_relaxed_***` to get closer to the actual PrEP
  numbers
- always use GFR RNG for OTC (unless `same` scenarios)
- always use rate based intervals for OTC PrEP?
- same for STI but with intervals for PrEP STD? (not sure) -> ask
<!-- NOTE: ask Sam -->

- To do:
    - adhr ratio: source?
    - for OTC: rate base HIV testing
    - always rate base GFR testing
    - graph of STD vs OTC: add IR100 HIV?
    - different best guess were:
        - std start rate stays the same, adjust with otc
        - same start rate but common OR
    - redo calibration?
